Stack Safety as a Security Property

### 2020-07-09 main.tex
Overhaul of section 7, discussion of lazy policies/properties, a few example fixes

\setcounter{pcctr}{27}
% H - Body
\vspace*{0.2em}
\[
\begin{array}{c|c|c|c}
  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
      {Set {\tt w} to 18}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
      {Deallocate {\tt h}'s call frame}
  \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
      {Load argument {\tt z} to $r_5$}
  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
      {Print {\tt z}}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
      {Add $r_4$ and $r_5$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
      {Store $r_4$ as the return value}
  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
      {Deallocate {\tt g}'s local state}
  \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 8}{Return to {\tt main} ($\jalr$)}
\end{array}
\]
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\fulloutc}[{\makebox[0pt]{Output}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{6}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{0}
\MemoryLabel{-47em}{0.75em}{42}
\MemoryLabel{-43em}{0.75em}{0}
\MemoryLabel{-39em}{0.75em}{0}
\MemoryLabel{-35em}{0.75em}{18}
\\
\end{center}
% Register File
\summary{8}{0}{\thepcctr}{104}{18}{0}
%\vspace*{0.2em}
\setcounter{pcctr}{8}
\[
\begin{array}{c|c|c|c}
  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
      {Set {\tt y} to the result}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
      {Deallocate {\tt g}'s call frame}
  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
      {Load {\tt x} to $r_5$}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
      {Add {\tt x} and {\tt y}}
  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
      {Print {\tt x + y}}
\end{array}
\]
\vspace*{0.2em}
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{31.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\fulloutc}[{\makebox[0pt]{Output}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
\memory{8}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{60}
\MemoryLabel{-47em}{0.75em}{60}
\MemoryLabel{-43em}{0.75em}{42}
\\
\end{center}
% Starting Register File
\summary{\thepcctr}{0}{37}{102}{60}{42}
\caption{Detailed Execution Trace
%  \rb{A suggestion for the description of $\jal$ and $\jalr$:
%    call(function label) [if labels are added next to the entrypoints]
%    or call(address of entrypoint) and return()?}
}
\label{fig:running-trace-b}
\end{figure}

@@

Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
\(\MPOTRACE\). We will say \(\MPT\) satisfies {\em trace integrity} with
respect to \(\contour\) if its final state (if any) agrees with its
initial one on all components marked \(\HIGHSEC\) in \(\contour\):
%
\[\begin{split}
  \propdef{TraceIntegrity} ~ \contour ~ \MPT \triangleq &
    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
    \mach_e(\component) = \mach_r(\component) \\
\end{split}\]
\sna{Experimenting with font on the mathy version of the property.}

@@

{\em Stack integrity} is a property of a system with respect to a call map
\(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
and \(\mpstate{\mach_e}{\pol_e}\) in every legal trace exhibits trace integrity
on the prefix up to its return, if any. For any initial state \(\mach_0\)
we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow
\MPT\). For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\), 
\(\propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
system as a whole enjoys stack integrity with respect to \(\callmap\).

@@

\[\begin{split}
  \propdef{TraceConfidentiality} ~ R ~ \MPT ~ \machT \triangleq
  & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
  & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land
    \mathit{FailCase} ~ R ~ \MPT ~ \machT \\
\end{split}\]

@@

If, for all such pairs of traces, \(\propdef{TraceConfidentialiy} ~
\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
stack confidentiality with respect to \(\callmap\).

@@

\section{Lazy Stack Safety}
\label{sec:lazy}

    \rb{Suggestion: name this ``observable \emph{trace? something else?}
      integrity'' to avoid confusion with observable stack integrity (in fact,
      it is called exactly this when discussing the example below). The rollback
      may also be a definition of its own to match the granularity of
      definitions in previous sections.}

@@

    \paragraph{Rolling Back to Idealized State}
      Observable integrity weakens eager integrity by allowing a returned
      state to differ from its call state at high integrity components,
      provided that the differences do not change the observable behavior
      of the rest of the program.  
%      provided the observable behavior of the induced trace from the return
%      matches that of an idealized state whose high integrity components do
%      match the call state.

      Given a contour, a call state, and a returned state, we define a
      ``rollback'' function \(\mathit{roll} : \CONTOURS \times \MACHS \times
      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
      the call state on those components that are high integrity in its
      contour, and matches the return state on low integrity components:
      \[\mathit{roll}(\contour,\mach_c,\mach_r)(\component) =
      \begin{cases}
        \mach_c(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
        \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
      \end{cases}\]

@@

      \apt{Much confusion here and below about whether $\mathit{roll}$ applies to call state
        or entry state just after.}

@@

      An MP-Trace enjoys {\em observable trace integrity} with respect to
      some contour \(\contour\) if its final state (if any) has observable
      behavior \apt{states don't have behavior} that prefixes that of its idealized version, constructed by
      \(\mathit{roll}\):

      \[\begin{split}
        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
        & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \\
        & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
          \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

@@

    However, if {\tt g} had an additional instruction to overwrite {\tt z}
    before printing, observable trace integrity would hold.
    \apt{Better to give an example in which {\tt h} overwrites something that
      is not needed to compute an output, so OTI does not fail.
      E.g., {\tt h} could overwrite {\tt w} with nonsense before returning -- but
      this doesn't matter, because the return value will replace the nonsense
      before this figures in the final print. 
      Also could be good to give an example in which we 
      overwrite {\tt x} instead of {\tt z} to show that we still fail OTI -- in other words,
      the entire remainder of the program matters, not just the caller.
    }

@@

    A system enjoys {\em observable stack integrity}
    % with respect to    \(\callmap\) and \(\codemap\)
    if, for any initial state \(\mach_0\) and
      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
      and contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).
\apt{more words please}

@@

    \paragraph{Restoring Varied States}

      Trace confidentiality is a property of two traces whose initial machine
      states are variations, with two sub-properties. First, the behavioral
      condition -- that their observable behaviors prior to the end of the
      trace are equivalent -- reflects the intuition that secret values should
      not influence visible behavior. Second, we have the end condition that
      their final states (if any) are equal up to the original variation. The
      end condition over-approximates the extension of the behavioral condition
      to future execution.\apt{That's quite mysterious.}

      We introduce {\em observable trace confidentiality}, which extends the
      behavioral condition to apply to execution from the final state on.
      This requires restoring varied components to their original values,
      since after return they will once again be legitimately accessible.

      To return to the example in \cref{fig:confex}, the varied final state
      \(\mach_r'\) might do anything depending on the values of the variation,
      so we cannot immediately compare traces from \(\mach_r\) and
      \(\mach_r'\). We must first restore the varied values to their original
      state, while preserving differences that have occurred during
      execution.
      \rb{To me, this reads like a prologue to the definitions properly
        speaking---maybe the definition header above should be a paragraph and
        a definition header appear next? I would also split it into two
        definitions.}
      \sna{Definitely better as a prologue. Where would you split it?}

%      In \cref{fig:lazyex1}, the final states of the original and variant
%      differ, because a written value (in black or white) depended on the value
%      of a high confidentiality location. So the resulting trace fails trace
%      confidentiality. For observable trace confidentiality, we ask whether the
%      differing values at address 42 cause any observable distinctions in
%      future behavior. To do so, we must undo the variation of formerly high
%      confidentiality locations, so that those in yellow match those in red,
%      while preserving the updated values where data may have been leaked.

@@

      Now we can describe a trace property that extends trace confidentiality.

@@

      Let \(\MPT\) and \(\machT\) be a pair of traces starting from
      \(\mpstate{\mach_e}{\pol_e}\) and \(\mach_e'\), respectively, with
      \(\mach_e \approx_\contour \mach_e'\). \apt{Should some of this move into the defining clauses below?} Then
      $\MPT$ and $\machT$ enjoy \emph{observable trace
        confidentiality} with respect to \(\contour\), written
            \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\), if:

      \begin{itemize}
      \item
      the two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
      \pi_o(\machT)\); and

      \item
      if \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\_}\), then \(\machT\) is finite, and
      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\), where
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\),
      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \hookrightarrow \machT'\),
      and \(\mach_r' = \last(\machT)\). 
      \end{itemize}

@@

@@

      A system enjoys {\em observable stack confidentiality}
      %with respect to \(\callmap\) and \(\codemap\)
      if, for any initial state
      \(\mach_0\) and each transition with entry state
      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).

>>>>

\setcounter{pcctr}{27}
% H - Body
\vspace*{0.2em}
\[
\begin{array}{c|c|c|c}
  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
      {Set {\tt w} to return value}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
      {Deallocate {\tt h}'s call frame}
  \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
      {Load argument {\tt z} to $r_5$}
  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
      {Print {\tt z}}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
      {Add $r_4$ and $r_5$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
      {Store $r_4$ as the return value}
  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
      {Deallocate {\tt g}'s local state}
  \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\PCname \leftarrow 8}{Return to {\tt main} ($\jalr$)}
\end{array}
\]
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\fulloutc}[{\makebox[0pt]{Output}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{6}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{0}
\MemoryLabel{-47em}{0.75em}{42}
\MemoryLabel{-43em}{0.75em}{0}
\MemoryLabel{-39em}{0.75em}{0}
\MemoryLabel{-35em}{0.75em}{18}
\\
\end{center}
% Register File
\summary{8}{0}{\thepcctr}{104}{18}{0}
%\vspace*{0.2em}
\setcounter{pcctr}{8}
\[
\begin{array}{c|c|c|c}
  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
      {Set {\tt y} to the result}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
      {Deallocate {\tt g}'s call frame}
  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
      {Load {\tt x} to $r_5$}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
      {Add {\tt x} and {\tt y}}
  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
      {Print {\tt x + y}}
\end{array}
\]
\vspace*{0.2em}
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{31.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\fulloutc}[{\makebox[0pt]{Output}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
\memory{8}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{60}
\MemoryLabel{-47em}{0.75em}{60}
\MemoryLabel{-43em}{0.75em}{42}
\\
\end{center}
% Starting Register File
\summary{\thepcctr}{0}{37}{102}{60}{42}
\caption{Detailed Execution Trace
%  \rb{A suggestion for the description of $\jal$ and $\jalr$:
%    call(function label) [if labels are added next to the entrypoints]
%    or call(address of entrypoint) and return()?}
}
\label{fig:running-trace-b}
\end{figure}

@@

Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
\(\MPOTRACE\). We will say \(\MPT\) satisfies {\em trace integrity} with
respect to \(\contour\) if its final state (if any) agrees with its
initial one on all components marked \(\HIGHSEC\) in \(\contour\):
%
\[\begin{split}
  \propdef{TraceInt} ~ \contour ~ \MPT \triangleq &
    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
    \mach_e(\component) = \mach_r(\component) \\
\end{split}\]
\sna{Experimenting with font on the mathy version of the property.}

@@

{\em Stack integrity} is a property of a system with respect to a call map
\(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
and \(\mpstate{\mach_e}{\pol_e}\) in every legal trace exhibits trace integrity
on the prefix up to its return, if any. For any initial state \(\mach_0\)
we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow
\MPT\). For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\), 
\(\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
system as a whole enjoys stack integrity with respect to \(\callmap\).

@@

\[\begin{split}
  \propdef{TraceConf} ~ R ~ \MPT ~ \machT \triangleq
  & \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \land \\
  & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land
    \mathit{FailCase} ~ R ~ \MPT ~ \machT \\
\end{split}\]

@@

If, for all such pairs of traces, \(\propdef{TraceConf} ~
\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
stack confidentiality with respect to \(\callmap\).

@@

\section{Stack Safety Properties for Lazy Enforcement}
\label{sec:lazy}

  A lazy policy permits some violations of integrity or confidentiality,
  justified because a truly harmful violation will be caught later.
  In a typical lazy version of the tag policy, writes are always permitted,
  but tagged to identify the context in which they were written. But reads
  check that the value read was written by the current context and fail-stop
  if it wasn't. So, if a callee makes an illegal write to its caller, when the
  caller tries to read it the policy will catch the error.

  The natural question is, how do we characterize the protections provided
  by these policies once stack safety is broken? How does a harmful violation
  that must be caught later differ from a harmless one? The properties in this
  section use the observation model to draw this distinction.

@@

    \paragraph{Rolling Back to Idealized State}
      Observable integrity weakens eager integrity by allowing a returned
      state to differ from its call state at high integrity components,
      provided that the differences do not change the observable behavior
      of the rest of the program. It is not immediately clear what constitutes
      changing behavior; we construct an idealized return state that contains
      only the changes that the caller was allowed to make, and use the trace
      from that state as a baseline against which the real trace will be
      compared.

      Given a contour, a call state, and a returned state, we define a
      {\em rollback} function \(\mathit{roll} : \CONTOURS \times \MACHS \times
      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
      the call state on those components that are high integrity in its
      contour, and matches the return state on low integrity components:
      \[\mathit{roll}(\contour,\mach_e,\mach_r)(\component) =
      \begin{cases}
        \mach_e(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
        \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
      \end{cases}\]

@@

@@

      A trace enjoys {\em observable trace integrity} with respect to
      some contour \(\contour\) if the trace induced by its final state
      (if any) has observable behavior that prefixes that of its idealized
      version, constructed by \(\mathit{roll}\):

      \[\begin{split}
        \propdef{ObsTraceInt} ~ \contour ~ \MPT \triangleq
        & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \\
        & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
          \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

@@

    However, consider instead if we insert after instruction 44 a new
    instruction \(45: [\rsp-4] \leftarrow 50\), overwriting {\tt w} in
    {\tt g}'s frame. So \(\mach_r'\) maps address 105 to 50, while the
    rolled-back state \(\mach_r'\) maps it to 0. But in the trace from either
    state, instructions 27 and 28 load the return value (18) and store it to
    105, eliminating the only difference. The remainder of the observation
    traces must be identical, so this trace does obey observable trace
    integrity.

@@

      A system enjoys {\em observable stack integrity}, like stack integrity,
      if, for any initial state and each \(n\)-transition from
      \(\mpstate{\mach_c}{\pol_c}\) to \(\mpstate{\mach_e}{\pol_e}\) in its
      induced trace, we take the prefix of the call
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{call} | \ret{\mach_c}\)
      and observable trace integrity holds on \(\MPT_{call}\) with
      contour \(\mathit{Cof}(\mach_c,n)\).

@@

      Two of the cases of trace confidentiality, \(\mathit{InfCase}\) and
      \(\mathit{FailCase}\), are purely properties of the observations produced
      by the traces they are properties of. If the traces are infinite, they
      should be similar; if one fail-stops, it should be a prefix. Only
      the \(\mathit{ReturnCase}\) checks machine states, because in that case
      the traces could continue, but have been cut off by a return. The key
      to making confidentiality lazy is to consider the traces after the
      return, considering only whether they are observationally similar
      or prefixes.

      This requires restoring varied components to their original values
      in the trace from the variant state, since after return they will
      once again be legitimately accessible. To return to the example in
      \cref{fig:confex}, the varied final state \(\mach_r'\) still has
      -5 as a return address for {\tt g}, so the trace after that return will
      behave oddly in a way that doesn't reflect a confidentiality violation.
      So we cannot immediately compare traces from \(\mach_r\) and
      \(\mach_r'\). We must restore the varied values while preserving
      differences that have occurred during execution.

@@

      Now we can describe a trace property akin to trace confidentiality, but
      continuing execution after the end of the callee trace.

@@

      {\em Observable trace confidentiality} is the lazy analog to trace
      confidentiality, and reuses the cases in which the traces don't return.
      For a contout \(\contour\), return predicate \(R\), and a pair of traces
      \(\MPT\) and \(\machT\), we consider the return case, in which
      \(\last(\MPT)\) is some \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and
      \(R ~ \mach_r\). Then \(\last(\machT)\) is some \((\mach_r',\obs_r')\)
      and \(R ~ \mach_r'\). As above, the observations of \(\MPT\) are similar
      to those of \(\machT\). In place of the requirement that \(\mach_r\)
      and \(\mach_r'\) agree on components, we have the following requirement.

      Finally, let \(\mach_e\) be the machine state of \(\head(\MPT)\),
      and \(\mach_e'\) be that of \(\head(\machT)\). We restore \(\mach_r'\)
      based on \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
      \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')\).
      The induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
      observational prefix of the induced trace from \(\mach_r'\),
      reflecting the possibility of a fail-stop after return.

      \[\begin{split}
        \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \triangleq
        & \head(\MPT) = \mpostate{\mach_e}{\pol_e}{\obs_e} \Rightarrow \\
        & \head(\machT) = (\mach_e',\obs_e') \Rightarrow \\
        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \last(\machT) = (\mach_r',\obs_r') \Rightarrow \\
        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mach_r'' \hookrightarrow \machT' \Rightarrow \\
        & \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
        & \pi_o(\MPT') \lesssim \pi_o(\machT') \\
      \end{split}\]

      We combine this case with the other two cases from ordinary trace
      confidentiality:

      \[\begin{split}
        \propdef{ObsTraceConf} ~ \contour ~ R ~ \MPT ~ \machT \triangleq
        & \mathit{ObsReturnCase} ~ \contour ~ R ~ \MPT ~ \machT \land \\
        & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
          \MPT ~ \machT \\
      \end{split}\]

@@

      On the other hand, suppose that we inserted an instruction at 43:
      \(r_6 \leftarrow [\rsp-8]\). \(r_6\) would hold 42 in \(\mach_r\)
      and -5 in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
      never used otherwise, the traces of {\tt h} will enjoy observable trace
      confidentiality.

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state and \(n\)-transition consisting of \(\mpstate{\mach_c}{\pol_c}\)
      and \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
      \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
      and \(\mach_e' \hookrightarrow \machT | \ret{\mach_c}\). A system
      enjoys observable stack confidentiality if, for all such
      \(\MPT\) and \(\machT\), \(\propdef{ObsTraceConf} ~
      \mathit{Cof}(\mach_c,n) ~ \ret{\mach_c} ~ \MPT ~ \machT\).
